Everything you wanted to know about PanDA but were truly truly afraid to ask

### 2017-03-04 PandaEvolution.tex
JEDI and double execution process

>>>>

\begin{itemize}
    \item Preserved logic by adapting PanDA Pilot to work as a (user land) persistent service instead of being submitted to the worknode. This was necessary because workernodes would not be able to communicate with PanDA Server.
    \item SAGA is used to submit jobs to the local batch system. Submitting jobs replaced spawning processes in the PanDA Pilot.
    \item Sergey provided architecture and reference paper.
\end{itemize}

### 2017-03-04 PandaOverview.tex
JEDI and double execution process

The design of the Production and Distributed Analysis (PanDA) system started in
2005 to develop a WMS for the ATLAS experiment. PanDA went into production for
the LHC run 1 between 2009 and 2013 and was then partially redesigned to be
deployed on run 2, between 2015 and 2018. New design features are constantly
prototyped and tested on small scale but, in this section, we summarize the
design, architecture, and execution process of the current state of the
practice.

\subsection{Design}

% \begin{itemize}
%   \item Serve both individual and production workflows
%   \item hide WLCG resources behind a single interface
%   \item queue-based interface to the application layer
%   \item pilot-based
%   \item job-based
%   \item global view of the properties and states of all the jobs
%   \item global view of the available resources as a single worldwide
%      processing resource.
%   \item centralize data management
%   \item X509 AAA
%   \item Multitenancy
% \end{itemize}

@@

PanDA's application model primarily assumes tasks, as single entities or
organized into workloads or workflows. Since 2005, a certain amount of
parallelism has been progressively introduced assuming both single core and
multi-threaded tasks~\cite{multithreaded_tasks} but, so far, no MPI tasks have
been considered for production. Tasks are assumed to have at least two types of
data dependences: one or more input file and one or more data sources. Task are
also supposed to be relatively self-contained, capable of setting up their own
execution environment or having a minimal set of common dependences. Different
types of task executable are assumed, both in terms of user-defined batch
scripts or application frameworks for group of users.

PanDA's user model is based on multitenancy and the support of at least two
types of HEP users: individual researchers and groups executing production-grade
workflows. Users are considered free to submit tasks and workflows at any point
in time, directly or via dedicated application frameworks. Consistently, PanDA's
security model is based on separation between authentication, authorization and
accounting for both single users and group of users. Both authentication and
authorization are based on digital certificates and the X.509 standard.

Currently, PanDA's execution model is based on five main abstractions: task,
job, queue, pilot, and event. Tasks are assumed to be partitioned into one or
more job, possibly of variable sizes. Jobs have both attributes and states and
are queued into a global queue for execution. Prioritization and binding of jobs
are assumed to depend on the attributes of each job. Pilot abstracts resource
capabilities and jobs are bound to pilots and executed by consuming their
resources. Consistently, pilots are considered to be multitenant as jobs
belonging to multiple workflows and/or users could be bound and executed on
them.

PanDA' data model uses events as the unit of data. Each event represents a
collision captured in the detector and one or more events can be packaged into
files or data source with other abstractions. As with jobs, data have both
attributes and states and some of the attributes are linked to those of jobs so
to create an association between the two otherwise separated entities. Data are
assumed to be stored centrally, replicated when necessary and transferred over
the network. This assumptions applies to both input and output data.

@@

\subsection{Architecture}

The implementation of PanDA consists of several interconnected subsystems, most
of them built from off-the-shelf and Open Source components. At a high level,
PanDA's architecture has six subsystems: PanDA Server, AutoPyFactory, PanDA
Pilot, PanDA Monitoring, JEDI, and Event Service. Subsystems communicate via
dedicated interfaces, implemented as API or http messaging. Databases are used
to store eventful entities like tasks, jobs and events, and to store information
about resources, logs, and accounting. Each subsystems is implemented as one or
more modules as shown in Figure~\ref{fig:architecture}. In the following we
briefly describe the overall architecture of PanDA and its components. The
description provides all the details required to understand how HPC resources
have been integrated within PanDA. For a full description of PanDA architecture
see~\cite{wiki, papers}.

Capabilities:
\begin{itemize}
  \item interoperable across job submission systems
  \item asynchronous distribution of input/out data organized as file collections associated to (set of) jobs.
  \item policy-aware scheduling
  \item Supporting multiple types of executable code.
\end{itemize}

% Components:
% \begin{itemize}
%   \item monitoring system
%   \item information system (static/dynamic information)
%   \item data management system
%   \item
% \end{itemize}

% \begin{itemize}
% \item PanDA Server
% \item Pilot
% \item Factory
% \item PanDA monitoring
% \item JEDI ??
% \item What else?
% \end{itemize}

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
  \end{center}
\caption{\mtnote{This diagram is a placeholder. I am working on a substitute
that will be referenced when describing PanDA's subsystems.}Schematic view of
the PanDA system\label{fig:architecture}, as deployed for production in 2017.
Black boxes indicate PanDA's subsystems; Orange boxes components of a subsystem;
purple boxes databases; green arrows http messaging interfaces; red arrows API;
purple arrows database communications.}
% Originally PanDA was designed for grid infrastructure... In this
% paper is focussed on the HPC use case; there are differences from
% the grid use case. This paper discusses how PanDA has been adapted
% to execute a workload % on HPC resources. The schematic overview is
% presented for workload X on Titan......
\end{figure}

\paragraph{\textbf{PanDA Server}} this subsystem is the central hub service of
PanDA and it operates as a web service. It runs on Apache web server,
interacting with the back-end database running on separate servers. It uses the
Apache worker model, with many independent  processes handling client requests
in parallel. Since all state is maintained in the central database, the PanDA
Server application instance itself is stateless. Currently production version of
PanDA utilizes Oracle database backend, but PanDA can also work with MySQL
family of databases.

@@

PanDA Server implementation consists of four components, all implemented in
Python and communicating via Python APIs: Task Buffer, Brokerage, Job
Dispatcher, and Data Service. Task Buffer is the management component of PanDA's
global queue and offers functionalities to: (i) store jobs into the global
queue; (ii) provide information about the state of each stored job; and (iii)
update the state of each job upon communication with the other three components
of PanDA Server. The Brokerage component enables (i) matching job attributes
with site and pilot attributes; (ii) managing the dispatch of input data to
processing sites; and (iii) deriving PanDA's data pre-placement requirement.
% The job/pilot matching algorithm assigns a weight to each candidate pilot and
% after applying a set of policies, the best ten candidates are taken.
The Job Dispatcher implements functionalities to: (i) receive requests from
active pilots for new jobs to process; (ii) retrieving from Task Buffer the jobs
with highest priority that can be executed on the resource on which the
requesting pilot has been instantiated; (iii) creating a wrapper for the
selected job, tailored to the resource of the requesting pilot; and (iv)
monitoring the state of the job execution on the pilot, updating the Task Buffer
component. Finally, the Data Service component offers functionalities to: (i)
retrieve information from Task Buffer and Brokerage about the data requirements
of each job and the resource to which each job has been assigned; (ii) enable
interaction with the ATLAS Distributed Data Management (DDM)~\cite{atlas_ddm} to
make input files available to the jobs and manage their output data.

\paragraph{\textbf{PanDA Pilot}} this subsystem is execution environment for
PanDA's jobs. Implemented in Python, a PanDA Pilot is launched by a wrapper
submitted to a target resource. Wrappers are tailored to the specific interfaces
and middleware exposed by each site so that PanDA Pilot remains
site-independent. PanDA Pilot implements four main functionalities: (i)
preparing the execution environment for PanDA's jobs, including recovering and
cleaning up previously failed jobs on that site and collecting information about
the compute, data, and network capabilities of the site; (ii) requesting a job
for execution to the Job Dispatcher component of PanDA Server; (iii) initiating
and monitoring the execution of pulled jobs, including setting up the job,
transferring its input when required, executing the payload, transferring
output, and checking log files and local space; and (iv) clean up when the
payload has finished to execute.

@@

PanDA Pilot implements the pilot paradigm~\cite{pilot_review} and, as such, it
decouples resource acquisition from job execution via multi-stage scheduling.
This has several benefits, including increasing the number of completed jobs per
unit of resource by lowering the time taken between finishing the execution of a
job an starting a the execution of a new job; enabling both concurrent and
sequential job execution depending on the number of cores available to the pilot
and the time for which the remain available; and exposing a unified and
consistent interface for the scheduling of jobs, independent from the type of
resource on which the pilot has been instantiated and the middleware installed
on the site. The specifics implementation of PanDA Pilot enables also better
reliability by real-time monitoring of the job execution process, and a certain
degree of fault-tolerance by recovering information and, possibly, output files
from jobs that failed, even before instantiating the PanDA Pilot.

@@

It should be noted that upon bootstrapping, a PanDA Pilot download major parts
of its runtime code a central Subversion repository via HTTP. The repository
works with an Apache web server, configured with a memory-based web proxy
(Squid). The purpose of the cache is to reduce the request load on the
back-end Subversion server. That improves performance \mtnote{Do we have
data/numbers?} since only occasional queries trigger a full lookup on the
back-end Subversion system, and most external queries are pulled from memory
on the front-end web server. The benefits of this system include a
high-performance code download service combined with code updates still being
immediately available as soon as they are committed to source code control.

\paragraph{\textbf{AutoPyFactory}} this subsystem implements a factory to submit
pilots locally and to both grid and cloud sites. Implemented to be executed as a
single demonized process, AutoPyFactory enable the concurrent spawning of
multiple job submission processes called APFQueue. Each APFQueue can communicate
with PanDA Server and the Brokerage module to collect information about the
state of the jobs it manages and, on the base of this information, submit new
pilots to one site per APFQueue. AutoPyFactory implements functionalities to
evaluate the number of pilots already instantiated on each site, and to apply a
consistent and adequate pressure on their batch queues. Submission is performed
via the Condor-G batch system~\cite{}, enabling submission to multiple type of
sites. AutoPyFactory exposes a dedicated monitoring service, feeding information
to the PanDA Monitoring subsystem. A proxy manager is also run, enabling
certificate-based authentication and authorization.

@@

\paragraph{\textbf{PanDA Monitoring}} this subsystem is implemented as a
web-based, dashboard-style graphical application. It runs on Apache and
interacts with the back-end database of PanDA Server to enable persistence.
PanDA Monitoring enables users and site administrators to gather information
about the status of current and past jobs, data movement, and pilot factory.
PanDA Monitoring allows users access to all job log files on job by job basis,
greatly simplifying code debugging and failure analysis in a distributed
computing environment.

\paragraph{JEDI} \mtnote{Should we include it?}

\paragraph{Event Service} \mtnote{Should we include it?}

\subsection{Execution Process}

\mtnote{Probably move this to the workflow}
\mtnote{Think whether a comparison with other WMS based on the review we did
makes sense in this context.}

\mtnote{Summary: JEDI retrieves task descriptions and source files, converting
them into job descriptions and events sets. Jobs are submitted to PanDA Server
while events to the Event Service. The PanDA Server schedules jobs to pilots
that have been submitted to resources via the PanDA Factory. Once executed, jobs
pulls events from the Event Service as input and push output back to JEDI.}

>>>>

The design of the Production and Distributed Analysis (PanDA) system started in
2005 to develop a WMS for the ATLAS experiment. PanDA went into production for
the LHC Run 1 between 2009 and 2013 and was then partially redesigned to be
deployed on Run 2, between 2015 and 2018. In this section, we summarize the
design, architecture, and execution process of the current state of the
practice.

% -----------------------------------------------------------------------------
\subsection{Design}
\label{ssec:panda_design}

@@

PanDA's application model primarily assumes single tasks, workloads and
workflows. Tasks represent a set of operations performed on a set of events,
stored across input files. Tasks can be decomposed into jobs, where each job
represents the task's set of operations and a partition of the task's events.
Since 2005, a certain amount of parallelism has been progressively introduced
for jobs~\cite{multithreaded_jobs} but, so far, no MPI jobs have been considered
for production. Jobs are supposed to be relatively self-contained, capable of
setting up their own execution environment or having a minimal set of common
dependences. Different types of job payload (i.e., executable) are assumed, both
in terms of user-defined batch scripts or application frameworks for group of
users.

PanDA's user model is based on multitenancy and the support of at least two
types of HEP users: individual researchers and groups executing production-grade
workflows. Users are considered free to submit tasks and workflows to the PanDA
WMS at any point in time, directly or via dedicated application frameworks.
Consistently, PanDA's security model is based on separation between
authentication, authorization and accounting for both single users and group of
users. Both authentication and authorization are based on digital certificates
and the X.509 standard and on the virtual organization abstraction.

Currently, PanDA's execution model is based on five main abstractions: task,
job, queue, pilot, and event. Both tasks and jobs are assumed to have attributes
and states and to be queued into a global queue for execution. Prioritization
and binding of jobs are assumed to depend on the attributes of each job. Pilot
is used to indicate the abstraction of resource capabilities. Jobs are thought
to be bound to pilots and the execution of their payload is performed on the
site where the pilots has been instantiated. Consistently, pilots are considered
to be multitenant as jobs belonging to multiple workflows and/or users can be
bound and executed on them.

PanDA' data model uses events as the unit of data. Each event represents a
collision captured in the detector and one or more events can be packaged into
files or data source with other container abstractions. As with jobs, data have
both attributes and states, and some of the attributes are shared between events
and jobs so to create associations. Data are assumed to be stored centrally,
replicated when necessary and transferred over the network. This assumptions
applies to both input and output data.

@@

% -----------------------------------------------------------------------------
\subsection{Architecture}
\label{ssec:panda_arch}

The implementation of PanDA consists of several interconnected subsystems, most
of them built from off-the-shelf and Open Source components. At a high level,
PanDA's architecture has six main subsystems: PanDA Server, AutoPyFactory, PanDA
Pilot, PanDA Monitoring, JEDI, and Event Service. Subsystems communicate via
dedicated API or http messaging. Databases are used to store eventful entities
like tasks, jobs and events, and to store information about resources, logs, and
accounting. Each subsystems is implemented by one or more modules as shown in
Figure~\ref{fig:architecture}. In the following, we briefly describe the overall
architecture of PanDA and its components. The description provides all the
details required to understand how HPC resources have been integrated within
PanDA (\ref{sec:integration}). For a more detailed description of PanDA
architecture see~\cite{wiki, papers}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
  \end{center}
\caption{\mtnote{This diagram is a placeholder. I am working on a substitute
that will be referenced when describing PanDA's subsystems.}Schematic view of
the PanDA system, as deployed for production in 2017. Black boxes indicate
PanDA's subsystems; Orange boxes components of a subsystem; purple boxes
databases; green arrows http messaging interfaces; red arrows API; purple arrows
database communications.}
% Originally PanDA was designed for grid infrastructure... In this
% paper is focussed on the HPC use case; there are differences from
% the grid use case. This paper discusses how PanDA has been adapted
% to execute a workload % on HPC resources. The schematic overview is
% presented for workload X on Titan......
\label{fig:architecture}
\end{figure}

\paragraph{\textbf{PanDA Server}} this subsystem is the central hub service of
PanDA and it operates as a web service. It runs on Apache web server,
interacting with the back-end database running on separate servers. It
implements the Apache worker model, with many independent processes handling
client requests in parallel. Since all state is maintained in the central
database, the PanDA Server application instance itself is stateless. Currently
production version of PanDA utilizes Oracle database backend, but PanDA can also
work with MySQL family of databases.

@@

PanDA Server implementation consists of four components, all implemented in
Python and communicating via Python APIs: Task Buffer, Brokerage, Job
Dispatcher, and Data Service. Task Buffer is the management component of PanDA's
global queue and offers functionalities to: (i) store jobs into the global
queue; (ii) provide information about and update the state of each stored job.
The Brokerage component enables (i) matching job attributes with site and pilot
attributes; (ii) managing the dispatch of input data to processing sites; and
(iii) deriving PanDA's data pre-placement requirement.
% The job/pilot matching algorithm assigns a weight to each candidate pilot and
% after applying a set of policies, the best ten candidates are taken.
The Job Dispatcher implements functionalities to: (i) receive requests from
active pilots for new jobs to process; (ii) retrieving from Task Buffer the jobs
with highest priority that can be executed on the resource on which the
requesting pilots have been instantiated; (iii) creating a wrapper for the
selected job, tailored to the resource of the requesting pilot; and (iv)
monitoring the state of the job execution on the pilot, updating the Task Buffer
component. Finally, the Data Service component offers functionalities to: (i)
retrieve information from Task Buffer and Brokerage about the data requirements
of each job and the resource to which each job has been assigned; (ii) enable
interaction with the ATLAS Distributed Data Management (DDM)~\cite{atlas_ddm} to
make input files available to the jobs and manage their output data.

\paragraph{\textbf{PanDA Pilot}} this subsystem implements the execution
environment for PanDA's jobs. Implemented in Python, a PanDA Pilot is launched
by a wrapper submitted to a site. Wrappers are tailored to the
specific interfaces and middleware exposed by each site so that PanDA Pilot
remains site-independent. PanDA Pilot implements four main functionalities: (i)
preparing the execution environment for PanDA's jobs,
% including  recovering and cleaning up previously failed jobs on that site and
collecting information about the compute, data, and network capabilities of the
site; (ii) requesting a job for execution to the Job Dispatcher component of
PanDA Server; (iii) initiating and monitoring the execution of pulled jobs,
including setting up the job, transferring its input when required, executing
the payload, transferring output, and checking log files and local space; and
(iv) clean up when the payload has finished to execute.

@@

PanDA Pilot implements the pilot paradigm~\cite{pilot_review}, decoupling
resource acquisition from job execution via multi-stage scheduling. This has
several benefits, including increasing throughput by lowering the time taken to
queue and schedule jobs; enabling both concurrent and sequential job execution
depending on the number of cores available to the pilot and the time for which
it remains available; and exposing a unified and consistent interface for the
scheduling of jobs, independent from the middleware of the site on which the
pilot has been instantiated.

% The specific implementation of PanDA Pilot promotes reliability by
% monitoring the execution of the job's payload, and a certain degree of
% fault-tolerance by recovering information and, possibly, output files from
% jobs that failed, even before instantiating the PanDA Pilot.

@@

It should be noted that upon bootstrapping, a PanDA Pilot download major parts
of its runtime code via HTTP from a central Subversion repository. The
repository works with an Apache web server, configured with a memory-based web
proxy (Squid). The purpose of the cache is to reduce the request load on the
back-end Subversion server. That improves performance \mtnote{Do we have
data/numbers?} since only occasional queries trigger a full lookup on the
back-end Subversion system, and most external queries are pulled from memory on
the front-end web server. The benefits of this system include a high-performance
code download service, combined with code updates still being immediately
available as soon as they are committed to source code control.

\paragraph{\textbf{AutoPyFactory}} this subsystem implements a factory to submit
pilots locally and to both grid and cloud sites. Implemented to be executed as a
single demonized process, AutoPyFactory enables the concurrent spawning of
multiple job submission processes called APFQueue. Each APFQueue can communicate
with PanDA Server and the Brokerage module to collect information about the
state of the jobs it manages and, on the base of this information, submit new
pilots to one site per APFQueue. AutoPyFactory implements functionalities to
evaluate the number of pilots already instantiated on each site, and to apply a
consistent and adequate pressure on their batch queues. Submission is performed
via the Condor-G batch system~\cite{condorg}, enabling submission to multiple
type of sites. AutoPyFactory exposes a dedicated monitoring service, feeding
information to the PanDA Monitoring subsystem. A proxy manager is also run,
enabling certificate-based authentication and authorization.

@@

\paragraph{\textbf{PanDA Monitoring}} this subsystem is implemented as a
web-based, dashboard-style graphical application. It runs on Apache and
interacts with the back-end database of PanDA Server to enable persistence.
PanDA Monitoring enables users and site administrators to gather information
about the status of current and past jobs, data movement, and pilot factory.
PanDA Monitoring allows users access to all log files of each job, greatly
simplifying code debugging and failure analysis in a distributed computing
environment.

\paragraph{\textbf{JEDI}} this subsystem implements a workload and workflow
management for tasks instead of jobs, as defined by HEP users. Implemented as a
standalone service that interacts with Panda Server via its REST interface, JEDI
offers two main functionalities: partitioning tasks to jobs, and binding jobs to
resources via dedicated queues. JEDI partition tasks to jobs by matching tasks
requirements to resource capabilities. This matching is based on tasks
attributes, and static and dynamic information about the resources of each
available site. Dynamic information is acquired via a small number of `scout'
jobs that collect information about current resources and their performance.
Functionally, JEDI defines a set of queues associated to groups of resources
with specific capabilities. Tasks are assigned to groups of these queues, and
the jobs in which each task is partitioned are assigned to each queue, depending
on their requirements. JEDI replaces the job brokerage functionalities of the
Brokerage module of the PanDA Server but this module still handles communication
between the Task Buffer and the Data Service modules.
%, using it to enact the execution of the jobs defined by JEDI
% on the resources and sites decided by JEDI.

% -----------------------------------------------------------------------------
\subsection{Management and Execution}
\label{ssec:panda_exec}

% The tasks of workloads and workflows submitted to PanDA are partitioned
% to one or more jobs. Each job is then executed on one of the available
% PanDA Pilots, chosen among those that the PanDA Server manages across multiple
% sites. Tasks processes a certain amount of events that are stored into files,
% and when a task is partitioned to multiple jobs, subsets of events need to be
% assigned to each job and made available to the pilot on which that job will be
% executed. Finally, the output of each job need to be retrieved and aggregated
% into the output of the original task (Fig.~\ref{fig:architecture}).

The relation between tasks and jobs can be one-to-one or one-to-many, and the
conversion between the two can by static or dynamic. During the LHC Run 1, PanDA
required users or the application layer to perform a static conversion between
tasks and jobs: tasks were described as a set of jobs with a fixed number of
events and then submitted to the PanDA Server (Fig.~\ref{fig:architecture}:A1).

This approach introduced inefficiency both with usability and resource
utilization. Ideally, users should not have to reason in terms of jobs: Users
conceive analyses in terms of one or more, possibly related tasks; the `job'
abstraction is required exclusively by the execution middleware, i.e. PanDA.
Further, a static partitioning of tasks into jobs does not take into account the
heterogeneity and dynamicity of the resources of the pilots on which each job
will be executed.

Another problem of static job sizing is that PanDA instantiates pilots on sites
with different type of resources and different models of availability of those
resources. An optimal sizing of each job should take into account these
properties. For example, sites may offer cores with different speed, networking
with different amount of bandwidth, and resources could be guaranteed to be
available for a defined amount of time or could disappear at any point in time
as with opportunistic models of resource provision.

JEDI was deployed for the LHC Run 2 to address these inefficiencies with a
machine-driven conversion of tasks into jobs. Users or the application layer (i)
submits tasks descriptions to JEDI; (ii) JEDI partitions tasks to jobs of
different size, depending on both static and dynamic information about available
resources; (iii) jobs are bound to sites with resources that best match jobs'
requirements; and (iv) jobs are submitted to the PanDA Server for execution
(Fig.~\ref{fig:architecture}:A2-B).

Once submitted to the PanDA Server, jobs are stored by the Task Buffer component
into a global queue implemented as a relational database
(Fig.~\ref{fig:architecture}:C). When jobs are submitted directly to the PanDA
Server, the Brokerage component is used to bind jobs to available sites,
depending on static information about the resources available for each site.
Jobs submitted by JEDI are instead already bound to resources so no further
brokerage is needed. Once jobs are bound to sites, the Brokerage module
communicates to the Data Service module what data sets need to be made available
on what site. The Data Service communicates these requirements to the ATLAS DDM
that, when needed, takes care of aggregating files into datasets and containers,
replicating them on the required sites.

The AutoPyFactory subsystem communicates with the Task Buffer component of the
PanDA Server, acquiring information about jobs that have been bound and are
ready for execution. On the base of this information, AutoPyFactory defines a
set of suitable PanDA Pilots, queues them into APFQueues, and concurrently
submits them to multiple sites.

% AutoPyFactory submits pilots only when similar pilots are not yet available
% and maintains a predefined amount of pressure on the queue of each site. In
% this way, resource availability and utilizations is optimized.

When PanDA Pilots become available on sites, they
% bootstrap, download their code, collect information about the execution
% environment and
call the Job Dispatcher module of the PanDA Server, requesting for a job to
execute \mtnote{Can a pilot request multiple jobs to execute concurrently?}. The
Job Dispatcher communicates with the Task Buffer, requesting for a job that is
bound to the site of that pilot and ready to be executed. Task Buffer checks the
global queue (i.e., the PanDA database) and if such a job is available, it
passes the job handler to the Job Dispatcher. The Job Dispatcher dispatches the
job to the PanDA Pilot.

Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
subprocess for the execution of the job's payload. The job is setup, input data
are transferred from the designated staging in location, and the job's payload
is executed. Once completed, output is transferred to the staging out location.
% During the job's payload execution, the monitoring process of PanDA Pilot
% checks the status of the job execution and its environment.

The Data Service module of the PanDA Server tracks and collects the output
generated by each job, updating jobs' attributes via the Task Buffer module.
When the output of all the jobs of a task are retrieved, they are aggregated and
made available to the user either via the PanDA Server of the JEDI service,
depending on how tasks or jobs were initially submitted.

\mtnote{Think whether a comparison with other WMS based on the review we did
makes sense in this context.}

### 2017-03-04 workflow.tex
JEDI and double execution process

The distinction between an abstract and concrete workflow instances highlights
an important element of the execution process of a workflow. Abstract workflows
can be interpreted as a formal description of the requirements for the workflow
execution, while concrete workflows can be interpreted as a formal description
of the workflow's execution process. The process of resource selection matches
the capabilities of resources to the requirements of a workflow, making possible
the transition from execution requirements to execution process.

As seen in \S\ref{sec:bkgrd}, Workload managers are those components of workflow
systems or, more in general, of distributed applications that enable resource
selection, acquisition, and use. The set of capabilities required by a workload
manager mostly depends on the heterogeneity and dynamicity of both the workflow
and the resources. Heterogeneity is a measure of the diversity of the workflow
requirements and resource capabilities; dynamicity is a measure of how much
these requirements and capabilities varies during execution.

The PanDA workload management system was initially designed to manage the
execution of static, relatively homogeneous tasks on dynamic and heterogeneous
resources~\cite{panda_grid}. Progressively, PanDA has evolved to enable the
management of more heterogeneous tasks and resources. In this paper, we describe
how the design and architecture of PanDA has enabled executions of
multi-threaded tasks on Titan, currently the largest high performance computing
(HPC) resource available in the USA to scientific research.

>>>>

% The distinction between an abstract and concrete workflow instances highlights
% an important element of the execution process of a workflow. Abstract
% workflows can be interpreted as a formal description of the requirements for
% the workflow execution, while concrete workflows can be interpreted as a
% formal description of the workflow's execution process. The process of
% resource selection matches the capabilities of resources to the requirements
% of a workflow, making possible the transition from execution requirements to
% execution process.

\begin{itemize}
    \item Is the Montecarlo Workflow a production workflow?
\end{itemize}

The design and implementation of PanDA enables the execution of diverse HEP workflows. Here we describe the execution process of the Monte Carlo workflow~\cite{montecarlo_workflow}, one of the largest workflows
executed by the ATLAS experiment

\subsection{NOTES - Please ignore}

As seen in \S\ref{sec:bkgrd}, Workload managers are those components of workflow
systems or, more in general, of distributed applications that enable resource
selection, acquisition, and use. The set of capabilities required by a workload
manager mostly depends on the heterogeneity and dynamicity of both the workload
and the resources. Heterogeneity is a measure of the diversity of the workflow
requirements and resource capabilities; dynamicity is a measure of how much
these requirements and capabilities varies during execution.

The PanDA workload management system was initially designed to manage the
execution of static, relatively homogeneous tasks on dynamic and heterogeneous
resources~\cite{panda_grid}. Progressively, PanDA has evolved to enable the
management of more heterogeneous tasks and resources.

% In this paper, we describe how the design and architecture of PanDA has
% enabled executions of multi-threaded tasks on Titan, currently the largest
% high performance computing (HPC) resource available in the USA to scientific
% research.
